Automated data-driven symbolic modeling of the lift forces acting on flapping wings

### 2021-02-07 paper/main.tex
Revisions to abstract, introduction and methods

\begin{abstract}
We explore a data-driven approach to symbolically modeling the lift produced by
the flapping wing of a micro air vehicle.  The aerodynamics of flapping flight
pose a modeling challenge due to complex effects related to vortex dynamics,
wake interaction, aeroelasticity and three-dimensional flow, yet simple and
accurate equations are useful for engineering purposes and optimization of wing
design.  Experiments were conducted to measure the flapping kinematics and
instantaneous lift forces produced by eleven different 3D printed wing
geometries at three different flapping amplitudes and multiple speeds, totaling
68 distinct sets of data.  58 of these data sets were used to train data-driven
models using a symbolic regression algorithm, while the remaining sets were
reserved for model testing.  Data-driven models were found to be more accurate
and much simpler than existing quasi-steady analytical models consistent with
aerodynamic theory.  Furthermore, data-driven models were found to separate
distinct physical effects into inertial and aerodynamic components bearing
remarkable resemblance to their analytical counterparts. This result suggests
that a data-driven approach could be a powerful tool in separating and
identifying the structural building blocks of a wide variety of unknown
dynamical systems.
\end{abstract}

@@

Forces produced by the flapping wings of insects, birds and micro air vehicles
are of crucial importance for flight, yet they are subtle and difficult to
model.  Many studies have developed equations and solution techniques to
understand the fluid flow around these wings, to compute control actions
required for stable flight, or to optimize wing design. Aerodynamic studies
related to insect flight and intermediate Reynolds numbers (Re 100-10,000) have
generally focused on either numerical solutions of Navier-Stokes equations in
two or three dimensions \cite{wu2004unsteady,wang2000two}, quasi-steady
approximations of the observed behavior
\cite{sane2002aerodynamic,ellington1984aerodynamics}, or a comparison of both
methods \cite{pesavento2004falling}.  The Navier-Stokes equations offer
important insights into the full character of the fluid around the wing
including vorticity and fluid-structure interaction, but are complex and
computationally expensive to calculate.  Quasi-steady approximations, in
contrast, capture most of the character of the aerodynamic forces without any
detailed modeling of the fluid flow.  These equations are phenomenological in
nature and constitute an assembly of terms that are consistent with aerodynamic
theory that can be used to represent lift, drag and other forces, as functions
of a wing's kinematic variables such as velocity and angle of attack.  Both
methods have shown qualitative and quantitative agreement with experimental
measurements \cite{wang2004unsteady}.

@@

Quasi-steady representations of lift are particularly well suited to the process
of symbolic regression, which harnesses computational power to conduct an
automated search over the space of all kinematic variables and geometric
parameters to find equations that most accurately predict the forces on a wing.
Symbolic regression algorithms are based on evolutionary computation
\cite{koza1992genetic,forrest1993genetic} and have been used to model complex
systems defined by explicit
\cite{duffy2002using,fillon2007symbolic,bautu2005symbolic}, implicit
\cite{schmidt2010symbolic}, iterated \cite{schmidt2009solving} and differential
equations \cite{bongard2007automated}.  The process underlying this evolutionary
computation approach is to generate an initially random population of equations
combining time-series data and constant variables with algebraic operators,
analytic functions and constants.  These candidate equations are then evaluated
to measure their accuracy in predicting the measured output of the system or
partial derivatives within the data.  In each iteration of the algorithm, the
best candidates are retained and varied by recombining and varying their
subexpressions. The algorithm is terminated when a desired level of accuracy is
reached.  In this case, the time-series data includes the orientations,
velocities and accelerations of wings, the constant variables include chord
lengths, wingspans, centroids of wing area and other physical constants, and the
known output of the system is the measured vertical force produced during
flapping.  The resulting equations returned by the algorithm are quasi-steady
models that can be compared to existing equations from the literature.

@@

\section*{Methods}
Our experimental apparatus consisted of a wing and actuator mounted to a
vertical force sensor, with a high-speed video camera recording the motion of
the wing from a top-down view and (using a mirror) also from a side-view. The
top-down and side views were used to detect and track the orientation of the
wing.  Figure~\ref{fig:apparatus} provides a diagram of the apparatus as well as
illustrations of the three primary variables describing the wing orientation:
Stroke ($\phi$), deflection ($\psi$) and deviation ($\theta$). The stroke motion
is enabled by a hinge joint and is the only actuated degree of freedom in the
system. The deflection of the wing is enabled by a hinge in the root of the
wing, and is a passive motion resulting from both inertial and aerodynamic
effects. Deviation refers to the passive vertical motion of the wing out of its
horizontal flapping plane, and is enabled only by the tolerances of the hinges
and the flexibility of the wing structure itself. These three variables, and
their derivatives, are some of the inputs required to model lift forces. The
other inputs are the geometric and inertial properties of the wings, obtained
from CAD designs and physical measurement. Next, we describe the wing design and
fabrication, the force measurement apparatus, and the method of extracting wing
kinematics from high-speed video.

@@

\begin{figure}
\centering
\includegraphics[trim=0 15 0 5, clip, width=0.5\textwidth]{figures/apparatus}
\caption{Schematic diagram of the experimental apparatus
  showing three kinematic variables of the wing motion: stroke ($\phi$),
  deflection ($\psi$) and deviation ($\theta$).\label{fig:apparatus}}
\end{figure}

\subsection*{3D Printed wings and actuation mechanism}
The experimental wings were fabricated using an Objet Connex500 3D printer with
the rigid Objet FullCure720 material.  Flapping flight with 3D printed wings was
reported in \cite{richter2011untethered}. The wing design consisted of an
interlocking root and tip, allowing the wing surface to pivot naturally during
flapping from its neutral vertical position to an angle of attack of
approximately 45\textdegree. The wing design is illustrated in
Figure~\ref{fig:wing1} and the hinge design is illustrated in
Figure~\ref{fig:root2}.  In total, eleven wings were used, ranging in wingspan
from 40 mm to 120 mm and ranging in chord length from 40 mm to 80 mm. Flapping
amplitudes were constrained by the geometry of the wing root design to be set at
72\textdegree, 81\textdegree, and 96\textdegree, and each wing was flapped at
speeds ranging between 1 Hz and 3.5 Hz for each of these amplitude settings.  68
sets of data were recorded, representing Reynolds numbers between 2500 and 5000.

@@

\subsection*{Measurement apparatus}
The apparatus used to measure the forces and kinematics of flapping wings was
based on a FUTEK LSM250 parallelogram load cell with a USB interface for data
logging.  This force measurement system was capable of a sampling rate of 1 kHz
and a maximum measurement precision of 1\e{-9} N.  Mounted directly to the load
cell was a wing-flapping mechanism, which consisted of a DC gear motor mated to
a crank and connecting rod used to drive the experimental wings in a roughly
sinusoidal flapping motion. This system is pictured in
Figures~\ref{fig:apparatus2} and \ref{fig:apparatus3}.  The natural frequency of
the load cell and wing driving mechanism was roughly 70 Hz, the natural
frequency for the first bending mode of the experimental wing was approximately
26 Hz, and experimental wings were flapped at or below approximately 3.5 Hz.
All experimental force measurements were filtered using a high-order zero-phase
digital low-pass filter with a cutoff frequency of 7.5 times the flapping
frequency, which preserved the primary inertial and aerodynamic forces acting on
the wings while eliminating the faster vibrations of the wing structure and the
load cell apparatus.

@@

\subsection*{High-speed measurement of flapping kinematics}
A Casio EXILIM EX-FH20 color video camera was used to record each experiment at
210 frames per second from a top-down perspective as well as a side view
provided by a mirror mounted in the field of view.  The main beam of each wing
was colored red to enable precise tracking of the stroke angle.  The top-down
projected area of the wing was used to extract deflection angle, and tracking of
the wing tip in the side-view was used to extract the angle of deviation out of
the wing's stroke plane.  These three angles precisely define the configuration
of the wing.  A schematic diagram of this apparatus is shown in
Figure~\ref{fig:apparatus}. A variety of image-processing techniques were used
to extract the kinematics of the wing motion from high-speed video.  The
red-colored main beam was identified with a color filter and a line was fitted
using a Hough Transform to identify the stroke angle.  The wing deflection angle
was computed by measuring the projected area of the wing in the image and
compared with the known dimensions of the rectangular wing shape. The
out-of-plane deviation angle of the wing was measured by tracking the leading
and trailing edge corners of the wingtip in the side-view of the mirror. We
refer the reader to our previous work for a more detailed description of our
methods of extracting wing kinematics from high speed video and processing the
measured data\cite{richter2012thesis}.

>>>>

\begin{abstract}
We explore a data-driven approach to symbolically modeling the lift produced by
flapping wings at the scale of a micro air vehicle.  The aerodynamics of
flapping flight pose a modeling challenge due to complex effects related to
vortex dynamics, wake interaction, aeroelasticity and three-dimensional flow,
yet simple and accurate equations are useful for engineering purposes and
optimization of wing design.  Experiments were conducted to measure the flapping
kinematics and instantaneous lift forces produced by eleven different 3D printed
wing geometries at three different flapping amplitudes and multiple speeds,
totaling 68 distinct sets of data.  58 of these data sets were used to train
data-driven models using a symbolic regression algorithm, while the remaining
sets were reserved for model testing.  Data-driven models were found to be more
accurate and much simpler than existing quasi-steady analytical models
consistent with aerodynamic theory.  Furthermore, data-driven models were found
to separate distinct physical effects into inertial and aerodynamic components
bearing remarkable resemblance to their analytical counterparts.  This result
suggests that a data-driven approach could be a powerful tool in separating and
identifying the structural building blocks of a wide variety of unknown
dynamical systems.
\end{abstract}

@@

Forces produced by the flapping wings of insects, birds and micro air vehicles
are of crucial importance for flight, yet they are subtle and difficult to
model.  Many studies have developed equations and solution techniques to
understand the fluid flow around these wings, to compute control actions
required for stable flight, or to optimize wing design. Aerodynamic studies
related to insect flight and intermediate Reynolds numbers (Re 100-10,000) have
generally focused on either numerical solutions of Navier-Stokes equations in
two or three dimensions \cite{wu2004unsteady,wang2000two}, quasi-steady
approximations of the observed behavior
\cite{sane2002aerodynamic,ellington1984aerodynamics}, or a comparison of both
methods \cite{pesavento2004falling}.  The Navier-Stokes equations offer
important insights into the full character of the fluid around the wing
including vorticity and fluid-structure interaction, but are complex and
computationally expensive to calculate.  Quasi-steady approximations, in
contrast, capture most of the character of the aerodynamic forces without any
detailed modeling of the fluid flow.  These equations are phenomenological in
nature and constitute an assembly of terms that are consistent with aerodynamic
theory that can be used to represent lift, drag and other forces, as functions
of a wing's kinematic variables such as velocity and angle of attack.  Both
methods have shown qualitative and quantitative agreement with experiments
\cite{wang2004unsteady}.

@@

Quasi-steady representations of lift are particularly well suited to the process
of symbolic regression, which harnesses computational power to conduct an
automated search over the space of relationships between kinematic variables and
geometric parameters to find equations that most accurately predict the forces
on a wing.  Symbolic regression algorithms are based on evolutionary computation
\cite{koza1992genetic,forrest1993genetic} and have been used to model complex
systems defined by explicit
\cite{duffy2002using,fillon2007symbolic,bautu2005symbolic}, implicit
\cite{schmidt2010symbolic}, iterated \cite{schmidt2009solving} and differential
equations \cite{bongard2007automated}.  The process underlying this evolutionary
computation approach is to generate an initially random population of equations
combining time-series data and constant variables with algebraic operators,
analytic functions and constants.  These candidate equations are then evaluated
to measure their accuracy in predicting the measured output of the system or
partial derivatives within the data.  In each iteration of the algorithm, the
best candidates are retained and varied by recombining and varying their
subexpressions. The algorithm is terminated when a desired level of accuracy is
reached.  In this case, the time-series data includes the orientations,
velocities and accelerations of wings, the constant variables include chord
lengths, wingspans, centroids of wing area and other physical constants, and the
known output of the system is the measured vertical force produced during
flapping.  The resulting equations returned by the algorithm are quasi-steady
models that can be compared to existing equations from the literature.

@@

\section*{Methods}
Our experimental apparatus consisted of a wing and actuator mounted to a
vertical force sensor, with a high-speed color video camera recording the motion
of the wing from a top-down view, and also from a side view using a mirror. The
top-down and side views were used to detect and track the position and
orientation of the wing.  Figure~\ref{fig:apparatus} provides a diagram of the
apparatus as well as illustrations of the three primary variables describing the
wing orientation: Stroke ($\phi$), deflection ($\psi$) and deviation ($\theta$).
The stroke motion is enabled by a hinge joint and is the only actuated degree of
freedom in the system. The deflection of the wing is enabled by a hinge in the
root of the wing, and is a passive motion resulting from both inertial and
aerodynamic effects. Deviation refers to the passive vertical motion of the wing
out of its horizontal flapping plane, and is enabled only by the tolerances of
the hinges and the flexibility of the wing structure itself. These three
variables, and their derivatives, are some of the inputs required to model lift
forces. The other inputs are the geometric and inertial properties of the wings,
obtained from CAD designs and physical measurement. Next, we describe the wing
design and fabrication, the force measurement apparatus, and the method of
extracting wing kinematics from high-speed video.

@@

\begin{figure}
\centering
\includegraphics[trim=0 15 0 5, clip, width=0.5\textwidth]{figures/apparatus}
\caption{Schematic diagram of the experimental apparatus
  showing three kinematic variables of the wing motion: Stroke ($\phi$),
  deflection ($\psi$) and deviation ($\theta$).\label{fig:apparatus}}
\end{figure}

\subsection*{3D Printed wings and actuation mechanism}
The experimental wings were fabricated using an Objet Connex500 3D printer with
the rigid Objet FullCure720 material. In prior work, we demonstrated flapping
flight of a micro air vehicle with wings made using this material
\cite{richter2011untethered}. The wing design consisted of an interlocking
hinged root and tip, allowing the wing surface to pivot naturally during
flapping from its neutral vertical position to a deflection angle of
approximately 45\textdegree.  The wing design is illustrated in
Figure~\ref{fig:wing1} and the hinge design is illustrated in
Figure~\ref{fig:root2}.  In total, eleven wings were used, ranging in wingspan
from 40 mm to 120 mm and ranging in chord length from 40 mm to 80 mm. Flapping
amplitudes were constrained by the geometry of the wing root design to be set at
72\textdegree, 81\textdegree, and 96\textdegree, and each wing was flapped at
speeds ranging between 1 Hz and 3.5 Hz for each of these amplitude settings.  68
sets of data were recorded, representing Reynolds numbers between 2500 and 5000.

@@

\subsection*{Measurement apparatus}
The force measurement apparatus was based on a FUTEK LSM250 parallelogram load
cell, capable of a sampling rate of 1 kHz and a precision of 1\e{-9} N.  Mounted
directly to the load cell was a wing-flapping mechanism, which consisted of a DC
gear motor powering a crank and connecting rod used to drive the experimental
wings in a roughly sinusoidal flapping motion. This system is pictured in
Figures~\ref{fig:apparatus2} and \ref{fig:apparatus3}.  The natural vibration
frequency of the load cell and wing driving mechanism was roughly 70 Hz, the
natural vibration frequency for the first bending mode of the experimental wing
was approximately 26 Hz, and experimental wings were flapped at or below
approximately 3.5 Hz.  All experimental force measurements were filtered using a
high-order zero-phase digital low-pass filter with a cutoff frequency of 7.5
times the flapping frequency, which preserved the primary inertial and
aerodynamic forces acting on the wings while eliminating the faster vibrations
of the wing structure and the load cell apparatus.

@@

\subsection*{High-speed measurement of flapping kinematics}
A Casio EXILIM EX-FH20 color video camera was used to record each experiment at
210 frames per second from a top-down perspective as well as a side view
provided by a mirror mounted in the field of view. Both the top-down and side
views featured matte black backdrops to facilitate easy identification of the
wing in video frames. A variety of image-processing techniques were used to
extract the kinematics of the wing motion from each video frame. The main beam
of each wing was colored red and identified with a color filter in the top-down
view. A line was then fitted using the Hough Transform to identify the stroke
angle. The wing deflection angle was computed by measuring the projected area of
the wing in the top-down view, calibrated using the known dimensions of the
rectangular wing shape. The out-of-plane deviation angle of the wing was
measured by tracking the leading and trailing edge corners of the wingtip in the
side view and comparing their positions to the nominal position of the wingtip
with zero deflection and deviation throughout the wing stroke. We refer the
reader to our previous work for further details and illustrations of our methods
for extracting wing kinematics from high speed video and processing the measured
data~\cite{richter2012thesis}.
